MIRSVM: Multi-Instance Support Vector Machine with Bag Representatives

### 2017-07-22 ArticleRevision/article/article.tex
Added references that Alberto had in this email and integrated them into the text.

One of the major complexities associated with MIL is the ambiguity of the relationship between a bag label and the instances within the bag~\citep{Alpaydin2015}. This stems from the standard MI assumption, where the underlying distribution among instances within positive bags is unknown. There have been different attempts to overcome this complexity, such as ``flattening" the MIL datasets, meaning instances contained in positive bags each adopt a positive label, allowing the use of classical supervised learning techniques~\citep{Ray2005}. This approach assumes that positive bags contain a significant number of positive instances, which may not be the case, causing the classifier to mislabel negative instances within the bag, decreasing the power of the MI model.

@@

Support vector machines represent a set of supervised, linear and nonlinear, classification and regression methods that have theoretical foundations on Vapnik-Chervonenkis (VC) theory~\citep{Huang2006,Schoelkopf2002}. SVM models are similar to other machine learning algorithms and techniques, but research has shown that they usually outperform them in terms of computational efficiency, scalability, and robustness against outliers~\citep{Kecman2016,Melki2016}, which makes them a useful data mining tool for various real-world applications.

@@

In addition to the SMI assumption, alternative MI assumptions have been proposed to date~\citep{Foulds2010}. A recent review describing the taxonomy of multi-instance classification was presented by~\citet{Amores2013}. The review presents various methods and algorithms used in literature which are categorized based on their approach to handling the MI input space. Instance-based classifiers that fall under the \textit{instance-space paradigm}, aim to separate instances in positive bags from those in negative ones. Bag-level classifiers (\textit{bag-space paradigm}) treat each bag as a whole entity, implicitly extracting information from each bag in order to accurately predict their labels. Methods that fall under the \textit{embedded-space paradigm} map the input bags to a single feature vector that explicitly encapsulates all the relevant information contained within each bag.

@@

SVMs are a particularly useful for learning linear predictors in high dimensional feature spaces, which is a computationally complex problem. In the context of classification, this problem is approached by searching for the optimal \textit{maximal margin} of separability between classes. Since most real-world datasets are non-linearly separable, ~\citet{Cortes1995} introduced the \textit{Soft-Margin L$1$-SVM}:
\begin{subequations} 
\label{eq:primalsvm}
\begin{alignat}{2}
\min\limits_{\bm w,b,\bm\xi}& {\,\,} \frac{1}{2}||\bm{w}||^2 + \frac{C}{m}\sum_{i=1}^m \xi_i \tag{\ref{eq:primalsvm}} \\ 
\text{s.t.}& {\,\,} y_i(\langle\bm{w},\bm{x}_i\rangle + b) \geq 1 - \xi_i,\,\, \forall i \in \{1,\ldots,m\}  \\
{} & {\,\,} \xi_i \geq 0,\,\, \forall i \in \{1,\ldots,m\},
\end{alignat}
\end{subequations} 
where $\bm w \in \reals^d$ is a $d$-dimensional weight vector, $b \in \reals$ is a bias term, and $C \in \reals$ is the penalty parameter that controls the trade-off between margin maximization and classification error minimization. The slack variable $\bm \xi \in \reals^m$ allows for optimizing over sample errors. However, if the training dataset is not linearly separable, the classifier may not have a good generalization capability~\citep{Nocedal2006}. Generalization and linear separability can be enhanced by mapping the original input space to a higher dimensional dot-product space by using a kernel function shown in Equation~\eqref{eqn:kerneltrick}:
\begin{equation}
\spa{K}\left(\bm{x}_i,\bm{x}_j\right) = \langle \phi\left(\bm{x}_i\right),\,\phi\left(\bm{x}_j\right)\rangle,
\label{eqn:kerneltrick}
\end{equation}
where $\phi\left(\cdot\right)$ represents a function mapping from the original feature space to a higher dimensional space. This kernel mapping is helpful when solving the dual SVM problem shown in Equation~\eqref{eq:dualsvm}.
\begin{subequations}
\label{eq:dualsvm}
\begin{alignat}{2}
\max\limits_{\bm \alpha} &\,\,\,\, \frac{1}{2}\sum_{i,j=1}^m{\alpha_i \alpha_jy_iy_j\spa{K}\left(\bm{x}_i,\bm{x}_j\right)} - \sum_{i=1}^m \alpha_i \tag{\ref{eq:dualsvm}} \\ 
\text{s.t.} &\,\,\,\, \sum_{i=1}^m \alpha_iy_i = 0,  \\
&\,\,\,\, 0 \leq \alpha_i \leq \frac{C}{m},\,\, \forall i \in \{1,\ldots,m\}. 
\end{alignat}
\end{subequations}

>>>>

One of the major complexities associated with MIL is the ambiguity of the relationship between a bag label and the instances within the bag~\citep{Alpaydin2015}. This stems from the standard MI assumption, where the underlying distribution among instances within positive bags is unknown~\citep{li2013multiple}. There have been different attempts to overcome this complexity, such as ``flattening" the MIL datasets, meaning instances contained in positive bags each adopt a positive label, allowing the use of classical supervised learning techniques~\citep{Ray2005}. This approach assumes that positive bags contain a significant number of positive instances, which may not be the case, causing the classifier to mislabel negative instances within the bag, decreasing the power of the MI model.

@@

Support vector machines represent a set of supervised, linear and nonlinear, classification and regression methods that have theoretical foundations on Vapnik-Chervonenkis (VC) theory~\citep{Huang2006,Schoelkopf2002}. SVM models are similar to other machine learning algorithms and techniques, but research has shown that they usually outperform them in terms of computational efficiency, scalability, and robustness against outliers~\citep{Kecman2016,Melki2016}, which makes them a useful data mining tool for various real-world applications~\citep{chen2016structural}.

@@

In addition to the SMI assumption, alternative MI assumptions have been proposed to date~\citep{Foulds2010}. A recent review describing the taxonomy of multi-instance classification was presented by~\citet{Amores2013}. The review presents various methods and algorithms used in literature which are categorized based on their approach to handling the MI input space. Instance-based classifiers that fall under the \textit{instance-space paradigm}, aim to separate instances in positive bags from those in negative ones. Bag-level classifiers (\textit{bag-space paradigm}) treat each bag as a whole entity, implicitly extracting information from each bag in order to accurately predict their labels~\citep{qiao2017diversified}. Methods that fall under the \textit{embedded-space paradigm} map the input bags to a single feature vector that explicitly encapsulates all the relevant information contained within each bag.

@@

SVMs are a particularly useful for learning linear predictors in high dimensional feature spaces, which is a computationally complex problem~\citep{wang2016hierarchical}. In the context of classification, this problem is approached by searching for the optimal \textit{maximal margin} of separability between classes~\citep{carrasco2015multi,fan2017mreklm}. Since most real-world datasets are non-linearly separable, ~\citet{Cortes1995} introduced the \textit{Soft-Margin L$1$-SVM}:
\begin{subequations} 
\label{eq:primalsvm}
\begin{alignat}{2}
\min\limits_{\bm w,b,\bm\xi}& {\,\,} \frac{1}{2}||\bm{w}||^2 + \frac{C}{m}\sum_{i=1}^m \xi_i \tag{\ref{eq:primalsvm}} \\ 
\text{s.t.}& {\,\,} y_i(\langle\bm{w},\bm{x}_i\rangle + b) \geq 1 - \xi_i,\,\, \forall i \in \{1,\ldots,m\}  \\
{} & {\,\,} \xi_i \geq 0,\,\, \forall i \in \{1,\ldots,m\},
\end{alignat}
\end{subequations} 
where $\bm w \in \reals^d$ is a $d$-dimensional weight vector, $b \in \reals$ is a bias term, and $C \in \reals$ is the penalty parameter that controls the trade-off between margin maximization and classification error minimization. The slack variable $\bm \xi \in \reals^m$ allows for optimizing over sample errors. However, if the training dataset is not linearly separable, the classifier may not have a good generalization capability~\citep{Nocedal2006}. Generalization and linear separability can be enhanced by mapping the original input space to a higher dimensional dot-product space by using a kernel function shown in Equation~\eqref{eqn:kerneltrick}:
\begin{equation}
\spa{K}\left(\bm{x}_i,\bm{x}_j\right) = \langle \phi\left(\bm{x}_i\right),\,\phi\left(\bm{x}_j\right)\rangle,
\label{eqn:kerneltrick}
\end{equation}
where $\phi\left(\cdot\right)$ represents a function mapping from the original feature space to a higher dimensional space. This kernel mapping is helpful when solving the dual SVM problem shown in Equation~\eqref{eq:dualsvm}.
\begin{subequations}
\label{eq:dualsvm}
\begin{alignat}{2}
\max\limits_{\bm \alpha} &\,\,\,\, \frac{1}{2}\sum_{i,j=1}^m{\alpha_i \alpha_jy_iy_j\spa{K}\left(\bm{x}_i,\bm{x}_j\right)} - \sum_{i=1}^m \alpha_i \tag{\ref{eq:dualsvm}} \\ 
\text{s.t.} &\,\,\,\, \sum_{i=1}^m \alpha_iy_i = 0,  \\
&\,\,\,\, 0 \leq \alpha_i \leq \frac{C}{m},\,\, \forall i \in \{1,\ldots,m\}. 
\end{alignat}
\end{subequations}
